\documentclass[letterpaper, 11pt]{article}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{float}
\usepackage{inputenc}
\usepackage[left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{extarrows}
\usepackage{xcolor}
\usepackage{lscape}
\usepackage{pdflscape}
\usepackage{pdfpages}
\usepackage{multicol}
\usepackage{leftindex}
\usepackage{algorithm2e}
\SetKwComment{Comment}{/* }{ */}
%\RestyleAlgo{ruled}
\usepackage{mathtools}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    }

% Listings
\usepackage{listings}
\usepackage{color}
\input{lst_style.tex}

% NewCommands
\newcommand{\peq}{ \mathrel{+}= }
\newcommand{\muleq}{ \mathrel{*}= }
\newcommand{\sign}{ \text{sign}}
\newcommand{\bm}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\mat}[1]{\begin{matrix} #1 \end{matrix}}
\newcommand{\lx}[2]{\leftindex #1 {#2}}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\itbf}[1]{\textit{\textbf{#1}}}
\newcommand{\mdet}[1]{\norm{\begin{matrix} #1 \end{matrix}}}
\newcommand{\pbox}[1]{\fbox{\begin{minipage}{\textwidth} #1 \end{minipage}}}
\newcommand{\lstln}[1]{\lstinline|#1|}
\newcommand{\lr}[1]{\left( #1 \right)}
\newcommand{\lrb}[1]{\left[ #1 \right]}


\title{Statement of Work - Purdue University \\ For Cummins Advanced Engineering Controls}
\author{
\itbf{Principal Investigator}: Prof. Peter H. Meckl, Purdue University}
\date{\today}


\begin{document}
\maketitle
\tableofcontents
\newpage
\
\newpage
%===============================================================================

\section{Objective}
To develop model-based non-intrusive on-board diagnostics (OBD) tools to detect catalyst-degradation in automotive Selective Catalytic Reduction (SCR) – Ammonia Slip Catalyst (ASC) system.

\section{Background}
Modern Diesel aftertreatment systems with Selective Catalytic Reduction (SCR) – Ammonia Slip Catalyst (ASC) need on-board diagnostics (OBD) tools that can accurately report SCR degradation level while avoiding false pass and false fail. The diagnostic tool’s performance must contain the following features: 1) high estimation accuracy; 2) high robustness against various noise factors under harsh on-road operating conditions; 3) minimal negative impact on SCR emissions control; and 4) cost-effective by using existing SCR configurations and commercially available NOx sensors.
Purdue University and Tennessee Technological University (TTU) have researched the literature on control-oriented SCR-aging models and various estimation methods for developing such a diagnostic tool. Since it is very difficult to imitate real-world catalyst degradation using test-cell accelerated aging, none of the existing control-oriented SCR-aging models have been validated or tested on real-world catalyst degradation data. Another limitation of existing OBD methods for SCR system is that they do not consider the presence of Ammonia Slip Catalyst (ASC) in series with SCR. Also, most existing OBD approaches cannot work under the limitations of commercial aftertreatment instrumentation such as no NH3 sensors, cross-sensitive NOx sensors etc. Therefore, development of better models to design OBD methods that can work with commercial systems under on-road conditions is required.
Purdue University and TTU have already initiated the research effort in collaboration with Cummins Inc. to fill this gap by investigating model-based and data-driven diagnostic algorithms to detect aging in SCR systems. The overall research endeavor includes: 1) achieving insight in SCR aging from modeling and diagnostic perspectives via the comprehensive analysis of experimental data; 2) developing accurate control-oriented aftertreatment models to predict reduced emissions performance at different degradation levels, and 3) developing robust and effective non-intrusive diagnostic methods that could distinguish SCR catalysts at different aging levels for real-world driving conditions.
As part of this effort, Cummins provided real-world truck data and test-cell data, along with seed funds for the year 2022, to Purdue University and TTU. In the year 2022, Purdue and TTU thoroughly studied the data, and explored several model-based and data-driven OBD approaches. For the model-based approach, Purdue developed a model and a model-based OBD for the operating conditions when ASC is active, and TTU has been working on a complementary model for conditions when the ASC is inactive. For the data-driven approach, Purdue worked on a binary classifier to classify the catalyst as either degreened (DG) or EUL at each time-stamp, and TTU has been working on exploring several methods to detect NOx-rich operating conditions which are suitable to extract distinct features from each truck. The motivation behind this proposal is to obtain funds for the year 2023 to continue this effort to develop OBD method capable of meeting the previously laid out requirements.
This is a joint effort by Purdue University and TTU and this document proposes Purdue’s share of the work for 2023. Purdue will focus on improving the model-based and data-driven OBD approaches with the following goals for 2023: 1) validate and improve the diagnostics-oriented SCR-ASC model using additional test-cell data and a high-fidelity SCR-ASC model from Cummins, 2) validate and improve both model-based and data-driven OBD methods using high-fidelity simulations and truck data with known aging levels (could be DG and/or EUL), and 3) support TTU in development and demonstration of model-based and data-driven OBD methods that rely on detecting NOx-rich operating conditions where ASC is inactive
Based on the findings from data analysis, models and diagnostic algorithms developed in 2022 by Purdue and TTU, the implementation and field validation of model-based non-intrusive diagnostic methods that could work with the Cummins existing commercial SCR systems will be planned in follow-on work.
Research Expertise and Qualification
The research team consists of Purdue University and TTU. The research expertise and qualification of Purdue University to successfully complete the project are described as follows:
The research group from Purdue University is led by Prof. Peter H. Meckl. Prof.
Meckl is a Professor of Mechanical Engineering at Purdue University. He received
his B.S. degree in Mechanical Engineering from Northwestern University in 1981,
M.S. in Mechanical Engineering from MIT in 1984, and Ph.D. in Mechanical
Engineering from MIT in 1988. He was the principal investigator for Purdue
EcoCAR2 project. He and his team have written 45 archival journal papers and
over 100 refereed conference papers, many of them concerning automotive control.
Much of the research on engine and aftertreatment diagnostics and control has
been sponsored by Cummins, work that spans over 25 years. He’s been involved in
research on modeling and control of urea-SCR systems starting in 2013 and has
graduated four M.S. thesis students who worked on modeling, estimation, and
control for the urea-SCR system. He has one PhD student currently
working on developing OBD for urea-SCR systems.


\section{Research Objectives}
\subsection{Objectives for 2023}
\begin{enumerate}
 \item In the current data, the operating conditions in the truck cover a wider range than the ones in test-cell. So, we wish to validate the SCR-ASC model on test-cell data with operating conditions similar to the truck data. If it is difficult to obtain such experimental data in test-cell, then use a high-fidelity SCR-ASC model in AVL Boost, to validate our low fidelity diagnostics-oriented model across a wide range of transient and steady operating conditions.
 \item Explore other low fidelity diagnostics-oriented SCR-ASC model structures, that could be calibrated using the signals between SCR and ASC that could be obtained from the high-fidelity Boost model.
 \item Validate the model-based and data-driven OBD results using truck data from trucks with DG and/or an EUL catalyst.
 \item Validate the model-based and data-driven OBD results using high-fidelity simulations.
 \item Based on the model and OBD validation results, iterate and improve both model-based and data-driven approaches to achieve robust diagnostics performance validated using high-fidelity simulations and actual data. Also implement a more rigorous and detailed version of the stochastic model-based OBD.
 \item Support TTU in development and demonstration of model-based, data-driven, and frequency-domain OBD methods that rely on detecting NOx-rich regions where ASC is inactive.
\end{enumerate}


\subsection{Objectives for 2024}
\begin{enumerate}
\item Identify the model paramters and parametric uncertainities using
the test-cell data and data from AVL boost model and validate them.
\item Develope the estimation metodology for the storage capacity (Observer Design).
\item Relate catalyst's aging to the storage capacity and develop robust
non-intrusive aging detection algorithm and test the algorithm on the test-cell
data and the road data segments.
\end{enumerate}
\subsection{Objectives post-2024}
\begin{enumerate}
    \item Provide an assessment of the potential to embed the algorithms in a real-time system.
    \item Implementation and comprehensive evaluation, including robustness analysis, of the proposed diagnostic algorithms in engine test cell and on the road.
\end{enumerate}



\section{Summary of work and results from 2023}
\subsection{Non-linear and Linearized Parametric Model}
We have the non-linear state-space model for SCR system consider only $NO$
dynamics among the NOx gasses and other related assumptions.

\begin{align*}
    \bm{x_1 \\ x_2 \\ x_3 \\ x_4} = \bm{C_{NO} \\ C_{NH_3} \\ M_{NH_3} \\ C_{NH_3, in}} \qquad &
    \bm{u_1 \\ u_2 } = \bm{C_{NO, in} \\ u_{inj}}
\end{align*}
\begin{align*}
    \mat{
    \\f_{13} &=& k_1
    \\f_{23} &=& k_{4F}
    \\f_{32} &=& k_{4F} \Theta
    \\f_{31} &=& k_1 V
    \\f_{24} &=& b_v F
    \\ b_b &=& \frac{1}{V}
    }
    \qquad
    \mat{
    \\ g_1    &=& b_v F
    \\ g_2    &=& b_v F + k_{4F} \Theta
    \\ g_{3}  &=& k_{4R}+k_3
    \\ g_{23} &=& k_{4R} V^{-1}
    \\ g_{32} &=& k_{4F} V \Theta
    \\ g_4 &=& \omega_u
    }
    \qquad
    \mat{
        b_{11} &=& b_v F
        \\
        b_{42} &=& \frac{\omega_u b_u}{F}
        \\ k_i &=& A_i e^{\lr{-E_a/RT}}
    }
\end{align*}
\begin{equation}{\label{eqn::ctrl_state}}
     \bm{\dot x_1 \\
        \dot x_2\\
        \dot x_3\\
        \dot x_4} =
    \bm{
        -f_{12} x_1 x_3
        -g_1 x_1
        \\
        %===
        -g_2 x_2
        + f_{23} x_2 x_3
        + g_{23} x_3
        + f_{24} x_4
        \\
        %===
        -f_{32} x_2 x_3
        -g_3 x_3
        -f_{31} x_1 x_3
        + g_{32} x_2
        \\
        %===
        -g_4 x_4
    }
    + \bm{b_{11} & 0\\
          0     & 0\\
          0     & 0\\
          0     & b_{42}  }\bm{u_1 \\ u_2 }
\end{equation}
\begin{equation}\label{eqn::ctrl_out}
    \bm{y_1 \\ y_2 \\ y_3} = \bm{1 & \chi & 0 & 0 \\
                                 0 & 1       & 0 & 0 \\
                                 0 & 0       & 0 & 0}
                            \bm{x_1 \\ x_2 \\ x_3 \\ x_4} +
                            \bm{0 & 0 \\
                                0 & 0 \\
                                1 & 0 \\}
                            \bm{u_1 \\ u_2}
\end{equation}


\subsection{Linearized Model}
Linearizing the above model w.r.t the states and temperature (T) and flow-rate (F).
\begin{align*}
     \bar f_{13} &= \bar k_{1}, \quad&
    \delta f_{13} &= \lr{ p_{1} \bar k_{1} } \\
    %===
     \bar f_{23} &= \bar k_{4F}, \quad &
    \delta f_{23}  &= \lr{ p_{4F} \bar k_{4F} } \\
    %===
     \bar f_{32} &= \bar k_{4F} \Theta, \quad &
    \delta f_{32}  &= \lr{ p_{32}  \bar k_{32} \Theta }
    \\
     \bar f_{31} &= \bar k_1 V, \quad &
    \delta f_{31}  &= \lr{ p_1 \bar k_1 V }
    \\
     \bar f_{24} &= b_v \bar F,
    \quad &
    \delta f_{24}  &= b_v
    \\
     \bar g_1 &= b_v \bar F,
    \quad &
    \delta g_1  &= b_v
    \\
    \bar g_2 &= b_v \bar F + \bar k_{4F} \Theta,
    \quad &
    \delta g_{2_F}  &= b_v
    \quad &
    \delta g_{2_T} &= \lr{ p_{32}  \bar k_{32}  \Theta }
    \\
    \bar g_3 &= \bar k_{4R} + \bar k_3,
    \quad &
    \delta g_3 &= \lr{ p_{4R} \bar k_{4R} + p_3 \bar k_3}
    \\
    \bar g_{23} &= \bar k_{4R} b_v,
    \quad &
    \delta g_{23}  &= \lr{ p_{4R} \bar k_{4R} b_v }
    \\
    \bar g_{32} &= \bar k_{4F} V \Theta,
    \quad &
    \delta g_{32}  &= \lr{ p_{4F} \bar k_{4F} V \Theta }
    \\
    g_4 &= \omega_u
    \\
    \bar b_{11} &= b_v \bar F,
    \quad &
    \delta b_{11} &= b_v
    \\
    \bar b_{42} &= \frac{\omega_u b_u}{\bar F},
    \quad &
    \delta b_{42}  &= -\lr{\frac{\omega_u b_u}{\bar F^2}}
\end{align*}


\begin{align*}
    \bm{\delta \dot x_1 \\ \delta \dot x_2 \\ \delta \dot x_3 \\ \delta \dot x_4} &= \bm{-\lr{ \bar f_{13} \bar x_3 + \bar g_1 } &
                    0 &
                    -\bar f_{13} \bar x_1 &
                    0 \\
                    %===
                    0 &
                    \lr{ -\bar g_2 + \bar f_{23} \bar x_3 }&
                    \lr{\bar f_{23} \bar x_2 + \bar g_{23}}&
                    \bar f_{24} \\
                    %===
                    -\bar f_{31} \bar x_3 &
                    \lr{-\bar f_{32} \bar x_3 + \bar g_{32} } &
                    \lr{-\bar f_{32} \bar x_2 - \bar g_3 - \bar f_{31} \bar x_1 } &
                    0 \\
                    %===
                    0 & 0 & 0 & -g_4
                    }
    \bm{\delta x_1 \\ \delta x_2 \\ \delta x_3 \\ \delta x_4}\\
            &\qquad+ \bm{ \bar b_{11} &
                            0 &
                            \delta f_{13} \bar x_1 \bar x_3 &
                            \lr{\delta b_{11} \bar u_1- \delta g_1 \bar x_1}
                            \\
                        %===
                        0&
                        0 &
                        \lr{-\delta g_{2_T} \bar x_2 + \delta f_{23} \bar x_2 \bar x_3 + \delta g_{23} \bar x_3} &
                        \lr{-\delta g_{2_F} \bar x_2 + \delta f_{24} \bar x_4 }
                        \\
                        %===
                        0&
                        0&
                        \lr{-\delta f_{32} \bar x_2 \bar x_3 - \delta g_3 \bar x_3 - \delta f_{31} \bar x_1 \bar x_3 + \delta g_{32} \bar x_2 } &
                        0
                        \\
                        %===
                        0&
                        \bar b_{42}&
                        0&
                        \delta b_{42} \bar u_2
                        }
    \bm{\delta u_1 \\ \delta u_2 \\ \delta T \\ \delta F}
\end{align*}
\begin{align*}
    \bm{\delta y_1 \\ \delta y_2 \\ \delta y_3} = \bm{1 & \chi & 0 & 0 \\
                                 0 & 1       & 0 & 0 \\
                                 0 & 0       & 0 & 0}
                            \bm{\delta x_1 \\ \delta x_2 \\ \delta x_3 \\ \delta x_4} +
                            \bm{0 & 0 & 0 & 0\\
                                0 & 0 & 0 & 0\\
                                1 & 0 & 0 & 0\\
                                }
                            \bm{\delta u_1 \\ \delta u_2 \\ \delta T \\ \delta F}
\end{align*}



\subsection{Ammonia Storage model and aging factor definition}
\input{secs/5_adsorption_model.tex}

\subsection{Cross-sensitivity Estimation}
\input{secs/7_cross_sensitivity_est.tex}

\section{Timeline for 2024}
\section{Deliverables}
\section{Other requirements}
\begin{itemize}
    \item Prior to the end of this project, a proposal for follow-on work will be submitted to Cummins.
    \item Cummins to provide expertise via mentorship and regular feedback.
    \item Cummins to provide funds to support the students.
    \item Cummins to provide test-cell data with operating conditions similar to on-road trucks
    \item Cummins to provide truck data with at least one and preferably two known aging levels (could be DG and/or EUL)
    \item Cummins to provide access to high-fidelity AVL Boost model for SCR-ASC. This could be a black-box model with access to pre-SCR, post-SCR, post-ASC signals, and NH3 storage values in SCR and ASC.
\end{itemize}

\section{Budget Information}
\newpage
\nocite{}
\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}
